Generalised gravitational burst generation with Generative Adversarial Networks

### 2020-05-29 Paper_draft/main.tex
results from cGAN

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
\bibliographystyle{iopart-num_custom}
\usepackage{xcolor}
\usepackage{pifont}
\usepackage{comment}
\usepackage{acronym}
\usepackage{standalone}
%\usepackage{import}
\usepackage{subfig}
\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing,angles,quotes,shapes.geometric, arrows,calc}
\expandafter\let\csname equation*\endcsname\relax 
\expandafter\let\csname endequation*\endcsname\relax 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\newcommand{\gguide}{{\it Preparing graphics for IOP Publishing journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}  
\newcommand{\jordan}[1]{\textbf{\textcolor{red}{JORDAN: #1}}}
\newcommand{\siong}[1]{\textbf{\textcolor{blue}{SIONG: #1}}}
\newcommand{\chris}[1]{\textbf{\textcolor{green}{CHRIS: #1}}}
\newcommand{\michael}[1]{\textbf{\textcolor{orange}{MICHAEL: #1}}}
\newcommand{\dcc}{LIGO-PXXXXXXX}
%\input{tag.tex}
\begin{document}

@@

\author{J. McGinn, C. Messenger, I.S. Heng, M. Williams}

@@

\section{Introduction}
\begin{comment}
\begin{itemize}
\item Need to introduce GWs - the current state of the field e.g. detections
and LVC papers \ding{51}
\item Introduce burst searches - what's the point of burst searches \ding{51} - lots of references 
\item Discuss the family of burst waveforms currently used and why - not in detail, just
an introduction \ding{51}
\item Introduce ML techniques in GWs \ding{51} - lots of references
\item What this paper does on GANs in 1 paragraph \ding{51}
\item Describe the structure of the paper 
\end{itemize}
\end{comment}

@@

% introduce burst searches
%
With these successes and continued upgrades to the detectors, further detections of CBCs are expected to be commonplace in a few short years. Another group of GW signals that has thus far been undetectable is GW ``bursts".  GW bursts are transient signals of typically short duration ($<$ 1s) whose
waveforms are not accurately modelled or are complex to re-produce.
Astrophysical sources for such transients include: Core collapse supernova,
Neutron star instabilities, Fallback accretion onto a neutron star, Non
axisymetric deformation in magnetars, Pulsar glitches and Neutron star
post-mergers.

@@

With the expectation that there will be many more \ac{GW} detections in the future, researches are starting to realise the need for fast and efficient \ac{GW} analysis to compete with the rising number of detections. While still in its infancy, Machine Learning (ML) with GW has already shown great potential for overcoming data bottlenecks. \ac{GW} analysis in areas of detection \cite{george2016deep,Gabbard2017,Gebhard_2019,Krastev_2020}, glitch classification \cite{Bahaadini, George_2018,Razzano_2018} and parameter estimation \cite{gabbard2019bayesian,shen2019deterministic,green2020gravitationalwave} have shown the success of in ML pattern recognition. Applying these algorithms show similar results to matched filtering techniques and have the benefit of being orders of magnitude faster.

@@

\begin{figure}
    \begin{minipage}[b]{0.3\linewidth}
        \subfloat[]{\label{fig:perceptron}\includegraphics[width=\linewidth,height=0.65\linewidth]{figures/neural_network-figure0.pdf}}
        \par
        \vspace{\baselineskip}
        \subfloat[]{\label{fig:tanh_activation}\includegraphics[width=\linewidth,height=0.65\linewidth]{figures/neural_network-figure1.pdf}}
    \end{minipage}
        \begin{minipage}[b]{0.7\linewidth}
        \subfloat[]{\label{fig:neural_network}\includegraphics[width=\linewidth,height=0.65\linewidth]{figures/neural_network-figure2.pdf}}
    \end{minipage}
    \caption{Neural Networks (a) A single neuron taking a vector of inputs and returning a singular output based on the weights, bias and activation function of the network. (b) The hyperbolic tangent used as an activation function. (c) A fully connected neural network containing two hidden layers that performs a mapping of an input vector to a singular output.}
\end{figure}

Since the birth of programmable computers, people have wondered if machines can develop true intelligence beyond formal mathematical procedures. Today, ML aims to learn apparent relationships held within the given data or `training data' in order to make accurate predictions without the need for additional programming . A common approach in ML relies on the computer learning on past experience to make decisions on future events. The hierarchical nature of this approach means that the machine can build complicated concepts from simpler ones. This branch of AI is called deep learning.
Neural networks are the quintessential machine learning algorithm that aims to approximate a function. They are built from many single processing units called neurons. The simpliest Neural Network is the perceptron layer Figure \ref{fig:perceptron} which holds a single neuron that takes several real inputs $x_{1}, ... , x_{i}$ and maps them to an output according to the following linear equation:

\begin{equation}
\sigma(\sum_i w_i x_i + b)
\label{eqn:neuron}
\end{equation}
where $w$ and $b$ are the weights and bias and $\sigma$ denotes the activation function. The weights are numbers which can be thought as the strength between connections. The output of a neuron is called its activation. Different choices of activation functions allow the user to simulate various models. A common example of an activation function is the hyperbolic tangent Figure \ref{fig:tanh_activation} that is used when the outputs of the network must be both positive and negative and re-scaled to $[-1,1]$. It is often useful to introduce a bias, $b$, such that the neuron remains inactive above zero but is active when the sum reaches a defined threshold. This bias is added before the activation function and it tells us how high the weighted sum needs to be before the neuron becomes active.

The output of a single neuron is defined by \ref{eqn:neuron} and gives a prediction $\hat{y}$ that can be compared to the real value $y$ through a cost or loss function. If the loss is non-zero, the network must work to minimise this function by updating the weights in the negative direction of the loss gradient in a process referred to as gradient descent. This loss function plays a critical role in how neural networks learn.

A neural network contains many single neurons connected in a layered structure as shown in Figure \ref{fig:neural_network}. The activations of the first layer (or input layer) act as the inputs to the second layer and so on until the output layer. Multilayered neural networks have intermediate layers between the input and output stages dubbed the hidden layers as the computations performed are not accessible to the user. The network can be thought of as a function F: $\mathbb{R}^N\rightarrow \mathbb{R}^M $ reliant on how activations from one layer bring activations in the next. The system is analogous to biology, where, some groups of neurons in animal brains cause certain others to fire.

@@

\begin{table}[hb]
\centering
\caption{Burst training parameters}
%\footnotesize
\begin{tabular}{@{} l l l l l l }
\br
\hline
 Waveform & Central frequency  & Decay & Central time epoch & Mass range \\
 & (Hz) & (s) & (s) & ($M_{\odot}$) \\
\mr
Sine-Gaussian & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & N/A  \\  
Ringdown & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & N/A \\
White-noise burst & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & N/A  \\
Gaussian pulse & N/A & 0.004 - 0.03 & 0.4 - 0.6 & N/A  \\
BBH & N/A & N/A & N/A & 5 - 70  \\
 \br
\end{tabular}\\
\label{Tab:training_parms}
\end{table}
\normalsize

GW burst signals remain an unmodelled phenomenon, as such, current
detection algorithms focus on waveforms signals appearing within multiple detectors We propose a
signal generation scheme using \acp{GAN} trained on burst-like
waveforms. BurstGAN is a conditional GAN that is trained on five signal morphology's spanning a range of prior
parameters. The parameter ranges can be seen in Table \ref{Tab:training_parms} The families are: sine-Gaussian $h(t) = A \exp\bigg[ - \frac{(t-t_{0})^2}{\tau^2} \bigg] \sin (2 \pi f_0 (t-t_0))$, a sinusoidal wave with a Gaussian envelop characterised by a central frequency, $f_0$; Ring-down, $h(t) = A \exp \bigg[-\frac{(t-t_0)}{\tau}\bigg]\sin(2 \pi f_0 (t-t_0))$, with frequency $f_0$ and duration $\tau$, white-noise bursts, with bounded frequency bandwidth $\Delta f$ and duration $\tau$, gaussian pulse $h(t) = \exp(-\frac{t^2}{\tau^2})$ with duration $\tau$ and binary-black hole inspirals which are simulated using IMRPhenomD waveform \cite{Khan_2016} routine from
LALSuite \cite{lalsuite} which models the
inspiral, merger and ringdown of a \ac{BBH} waveform. The component masses lie
in the range of [5,70]$M_{\odot}$ with zero spins
and we fix m$_1$ $>$ m$_2$. The mass distribution is approximated by a power
law with index of 1.6 \cite{Abbott_2019}. The signals are generated using random right ascensions and
declinations uniform over the sky and the inclinations are drawn from the
cosine of a uniform distribution in the range [-1,1]. The peaks of the
waveforms are set to be within [0.4,0.6]s of the 1s time interval.

@@

% specifically define our architecture
%
We adopt the suggestions of \cite{Radford2015,DBLP:journals/corr/abs-1809-11096}, lengthening
one-dimensional convolution kernels on both the generator and discriminator.
The Generator model is fully convolutional, upsampled using strided transposed
convolutions with batch normalisation in the first layer and ReLU activations
throughout with the exception of Tanh for the output layer. Each transposed
convolutional layer uses a kernel size of $18\times 1$ and stride of 2. The
discriminator network mirrors that of the generator without batch
normalization, using LeakyReLU activations, SpatialDropout, and a 2-stride
convolution for downsampling. The discriminator has two output layers: the
first output is a single node activated by a Sigmoid that can be interpreted as
the realness of the the signal, the second output is 5 nodes activated using
the softmax function predicting the class of the input.  This model is trained
with binary cross entropy for the first output and sparse categorical
cross-entropy for the second output. The full architecture description can be seen in Table \ref{Tab:hyperparameters}.~\chris{OK, very technically tight but too
computer sciencey. Try to make it human readbale for a physicist and refer to
the table of parameters in the appendix.}

% describe the hyperparameter tunings
%
Neural networks and subsequently \acp{GAN} have multiple parameters a developer
can tune when designing the model and these are referred to as hyperparameters.
The final network design used in this work comes from the use of trial and
error and the initial designs influenced by the available literature. After
tuning the multiple hyperparamters (Table~\ref{Tab:hyperparameters}), the GAN
was trained for $6\times 10^5$ iterations~\chris{what is an iteration here?
It's not an epoch right?} and takes $\Or$(1) day to train~\chris{trained on
what? Were there any major decisions made based on the trial and error process?
Anything that was learned from your mistakes that might be useful to others?}.

@@

% explain embedding
%
A learned embedding is an efficient way of representing categorical or class
based data. In contrast to traditional "one-hot encoding" where each class is
represented by a binary sparse vector, embeddings map each class to its own
distinct vector of a given size. The elements of these vectors are initialized
randomly and tuned during training just like weights~\chris{I thought that they
weren't trained. Are they definitely trained? Is this not just a single fully
connected layer that turns 5 numbers into 120? How is it any different to
that?}. This reduces computational cost as the sparse vectors containing mostly
zeros are not needed and allows related classes to cluster together. Once the
networks are trained the learned embedding vectors can be extracted from the
model and used as inputs to the generator. The benefit here is that the former
integer class labels are replaced by higher dimensional vector representations
allowing for finer experimenting in other applications. The class label is
interpreted as an additional channel early in the generator model. This is
achieved by projecting the class inputs as a learned embedding layer, into a
fully connected layer or "dense" layer which can then be concatenated
channel-wise to $\textbf{z}$.~\chris{very nice but still a bit too technical
without consideration for lowly astrophysicists. Can you soften it a bit to
make it more accessible. Explain more of the concepts when you introduce them.}

@@

\subsection{Waveform quality}
The generator network is a function G : $\mathbf{z},\mathbf{c},\mathbb{\textbf{s}}$ $\in$ $\mathbb{R}^{100}$ $\to$ $\mathbb{R}^{1024\times2}$, where $\mathbf{z},\mathbf{c},\mathbb{\textbf{s}}$ are the latent vector, class embedding vector and sky positions respectively. Given a latent vector randomly sampled from a normal distribution with zero mean and unit variance, a class label which is represented by a 120 dimensional vector for each class and antenna responses, the results from the generator can be seen in Figure \ref{fig:gen_signals}. Depending on the orientation of the detector with respect to a hypothetical signal in the sky, the waveforms may appear inverted, shifted in time and their strain attenuated. Each plot shows the output of the generator after given randomised $\mathbf{z}, \mathbb{\textbf{s}}$ and one of the five class vectors $\mathbf{c}$.

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{figures/conditional_gens.pdf}
    \caption{Generated waveforms after training and conditioning on five classes. The generator will output two waveforms as seen by detectors in Hanford (red) and Livingston (blue). The generator is able to capture the characteristics of each waveform and structure the class space to give control over which waveform to generate. Each row shows a random assortment of one of the five classes the GAN is trained on. }
    \label{fig:gen_signals}
\end{figure}

@@

\subsection{Latent space interpolation}
In this section we explore the latent space formed by the generator by interpolation. We take two random points in the latent space and linearly interpolate between them (six times inclusive). These new latent space vectors can now be fed into the generator to make predictions on while keeping the class vectors constant. The third input, the response values are kept constant for each class. The full effect shown in Figure \ref{fig:z_interp}. We can see that each plot shows plausible waveforms suggesting that the generator has constructed a smooth space unlike the discrete training case. Additionally as each class is given the same latent points to interpolate over, we can see that the waveforms cluster together with respect to their parameters. Visually, the sine-gaussian and ring-down waveforms share similar frequencies and the other signals show similar decays and starting epochs. The only exception is BBH waveforms, which is expected as they were trained with more variety of parameters and consistently have their peaks in the last quarter of the time series.

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{figures/z_interp_fix_c.pdf}
    \caption{Interpolations between two random latent points in z. Each row uniformly interpolates between two points in $\mathbf{z}$ keeping the class fixed. Only a single waveform from the generator is plotted and each signal is re-scaled to [-1,1] effectively removing the antenna responses for clarity.}
    \label{fig:z_interp}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/4_c_interp.pdf}
    \caption{Class space interpolation with latent space held constant throughout. The plots show a zoomed in section of the 1s time interval. Top row: Sine-Gaussian class to ringdown class, Middle Row: Sine-Gaussian class to whitenoise burst class, Bottom row: Sine-Gaussian class to Gaussian class.}
    \label{fig:c_interp}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/sg_bbh_interp.pdf}
    \caption{Class space interpolation with latent space held constant throughout. The full 1s interval is plotted for class based interpolations between a Sine-Gaussian and a BBH in spiral.}
    \label{fig:sgbbh_interp}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
\bibliographystyle{iopart-num_custom}
\usepackage{xcolor}
\usepackage{pifont}
\usepackage{comment}
\usepackage{acronym}
\usepackage{standalone}
\usepackage{academicons}
\usepackage{scalerel}
\usepackage{layouts}
%\usepackage{import}
\usepackage{subfig}
\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing,angles,quotes,shapes.geometric, arrows,calc}
\usetikzlibrary{svg.path}
\expandafter\let\csname equation*\endcsname\relax 
\expandafter\let\csname endequation*\endcsname\relax 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage[capitalize]{cleveref}
\newcommand{\gguide}{{\it Preparing graphics for IOP Publishing journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}  
\newcommand{\jordan}[1]{\textbf{\textcolor{red}{JORDAN: #1}}}
\newcommand{\siong}[1]{\textbf{\textcolor{blue}{SIONG: #1}}}
\newcommand{\chris}[1]{\textbf{\textcolor{green}{CHRIS: #1}}}
\newcommand{\michael}[1]{\textbf{\textcolor{orange}{MICHAEL: #1}}}
\newcommand{\dcc}{LIGO-PXXXXXXX}
%\input{tag.tex}
% Command for ORCID id's
\definecolor{orcidlogocol}{HTML}{A6CE39}
\tikzset{
  orcidlogo/.pic={
    \fill[orcidlogocol] svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3,0,128,0C198.7,0,256,57.3,256,128z};
    \fill[white] svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
                 svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
                 svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
  }
}

\newcommand\orcidicon[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
\begin{tikzpicture}[yscale=-1,transform shape]
\pic{orcidlogo};
\end{tikzpicture}
}{|}}}}

\usepackage{hyperref} %<--- Load after everything else

\begin{document}

@@

\author{
    J. McGinn \orcidicon{0000-0000-0000-0000},
    C. Messenger \orcidicon{0000-0001-7488-5022},
    I.S. Heng \orcidicon{0000-0000-0000-0000},
    M. J. Williams \orcidicon{0000-0003-2198-2974}
}

@@

\section{Introduction}
textwidth in inches: \printinunitsof{in}\prntlen{\textwidth}
\begin{comment}
\begin{itemize}
\item Need to introduce GWs - the current state of the field e.g. detections
and LVC papers \ding{51}
\item Introduce burst searches - what's the point of burst searches \ding{51} - lots of references 
\item Discuss the family of burst waveforms currently used and why - not in detail, just
an introduction \ding{51}
\item Introduce ML techniques in GWs \ding{51} - lots of references
\item What this paper does on GANs in 1 paragraph \ding{51}
\item Describe the structure of the paper 
\end{itemize}
\end{comment}

@@

% introduce burst searches
%
With these successes and continued upgrades to the detectors, further detections of CBCs are expected to be commonplace in a few short years. Another group of GW signals that has thus far been undetectable is GW ``bursts".  GW bursts are transient signals of typically short duration ($<$ 1s) whose
waveforms are not accurately modelled or are complex to re-produce.
Astrophysical sources for such transients include: Core collapse supernova,
Neutron star instabilities, Fallback accretion onto a neutron star, Non
axisymetric deformation in magnetars, Pulsar glitches and Neutron star
post-mergers. \michael{Might be worth citing some sources for this statement}

@@

With the expectation that there will be many more \ac{GW} detections in the future, researches are starting to realise the need for fast and efficient \ac{GW} analysis to compete with the rising number of detections. While still in its infancy, Machine Learning (ML) with GW has already shown great potential for overcoming data bottlenecks. \ac{GW} analysis in areas of detection \cite{Gabbard2017,Gebhard_2019,Krastev_2020}, glitch classification \cite{Bahaadini, George_2018,Razzano_2018} and parameter estimation \cite{gabbard2019bayesian,shen2019deterministic,green2020gravitationalwave} have shown the success of in ML pattern recognition. Applying these algorithms show similar results to matched filtering techniques and have the benefit of being orders of magnitude faster.

@@

\begin{figure}
    \begin{minipage}[b]{0.3\linewidth}
        \subfloat[]{\label{fig:perceptron}\includegraphics[width=\linewidth,height=0.65\linewidth]{figures/neural_network-figure0.pdf}}
        \par
        \vspace{\baselineskip}
        \subfloat[]{\label{fig:tanh_activation}\includegraphics[width=\linewidth,height=0.65\linewidth]{figures/neural_network-figure1.pdf}}
    \end{minipage}
        \begin{minipage}[b]{0.7\linewidth}
        \subfloat[]{\label{fig:neural_network}\includegraphics[width=\linewidth,height=0.65\linewidth]{figures/neural_network-figure2.pdf}}
    \end{minipage}
    \caption{Neural Networks (a) A single neuron taking a vector of inputs and returning a singular output based on the weights, bias and activation function of the network. (b) The hyperbolic tangent used as an activation function. (c) A fully connected neural network containing two hidden layers that performs a mapping of an input vector to a singular output. \michael{Equation in the neuron looks very small}}
\end{figure}

Since the birth of programmable computers, people have wondered if machines can develop true intelligence beyond formal mathematical procedures. Today, ML aims to learn apparent relationships held within the given data or `training data' in order to make accurate predictions without the need for additional programming . A common approach in ML relies on the computer learning on past experience to make decisions on future events. The hierarchical nature of this approach means that the machine can build complicated concepts from simpler ones. This branch of AI is called deep learning.
Neural networks are the quintessential machine learning algorithm that aims to approximate a function. They are built from many single processing units called neurons. The simpliest Neural Network is the perceptron layer \cref{fig:perceptron} which holds a single neuron that takes several real inputs $x_{1}, ... , x_{i}$ and maps them to an output according to the following linear equation:

\begin{equation}
\sigma(\sum_i w_i x_i + b)
\label{eqn:neuron}
\end{equation}
where $w$ and $b$ are the weights and bias and $\sigma$ denotes the activation function. The weights are numbers which can be thought as the strength between connections. The output of a neuron is called its activation. Different choices of activation functions allow the user to simulate various models. A common example of an activation function is the hyperbolic tangent \cref{fig:tanh_activation} that is used when the outputs of the network must be both positive and negative and re-scaled to $[-1,1]$. It is often useful to introduce a bias, $b$, such that the neuron remains inactive above zero but is active when the sum reaches a defined threshold. This bias is added before the activation function and it tells us how high the weighted sum needs to be before the neuron becomes active.

The output of a single neuron is defined by \cref{eqn:neuron} and gives a prediction $\hat{y}$ that can be compared to the real value $y$ through a cost or loss function. If the loss is non-zero, the network must work to minimise this function by updating the weights in the negative direction of the loss gradient in a process referred to as gradient descent. This loss function plays a critical role in how neural networks learn.

A neural network contains many single neurons connected in a layered structure as shown in \cref{fig:neural_network}. The activations of the first layer (or input layer) act as the inputs to the second layer and so on until the output layer. Multilayered neural networks have intermediate layers between the input and output stages dubbed the hidden layers as the computations performed are not accessible to the user. The network can be thought of as a function F: $\mathbb{R}^N\rightarrow \mathbb{R}^M $ reliant on how activations from one layer bring activations in the next. The system is analogous to biology, where, some groups of neurons in animal brains cause certain others to fire.

@@

\begin{table}[hb]
\centering
\caption{Burst training parameters}
%\footnotesize
\begin{tabular}{@{} l l l l l l }
\br
\hline
 Waveform & Central frequency  & Decay & Central time epoch & Mass range \\
 & (Hz) & (s) & (s) & ($\textrm{M}_{\odot}$) \\
\mr
Sine-Gaussian & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & N/A  \\  
Ringdown & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & N/A \\
White-noise burst & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & N/A  \\
Gaussian pulse & N/A & 0.004 - 0.03 & 0.4 - 0.6 & N/A  \\
BBH & N/A & N/A & N/A & 5 - 70  \\
 \br
\end{tabular}\\
\label{Tab:training_parms}
\end{table}
\normalsize

GW burst signals remain an unmodelled phenomenon, as such, current
detection algorithms focus on waveforms signals appearing within multiple detectors We propose a
signal generation scheme using \acp{GAN} trained on burst-like
waveforms. BurstGAN is a conditional GAN that is trained on five signal morphology's spanning a range of prior
parameters. The parameter ranges can be seen in \cref{Tab:training_parms} The families are: sine-Gaussian $h(t) = A \exp\left[ - (t-t_{0})^2 / \tau^2 \right] \sin (2 \pi f_0 (t-t_0))$, a sinusoidal wave with a Gaussian envelop characterised by a central frequency, $f_0$; Ring-down, $h(t) = A \exp \left[-{(t-t_0)} / {\tau} \right] \sin(2 \pi f_0 (t-t_0))$, with frequency $f_0$ and duration $\tau$, white-noise bursts, with bounded frequency bandwidth $\Delta f$ and duration $\tau$, gaussian pulse $h(t) = \exp(-t^2 / \tau^2)$ with duration $\tau$ and binary-black hole inspirals which are simulated using IMRPhenomD waveform \cite{Khan_2016} routine from
LALSuite \cite{lalsuite} which models the
inspiral, merger and ringdown of a \ac{BBH} waveform. The component masses lie
in the range of [5,70] $\textrm{M}_{\odot}$ with zero spins
and we fix m$_1$ $>$ m$_2$ \michael{I think the exponentials look neater without using frac\{\}\{\}}. The mass distribution is approximated by a power
law with index of 1.6 \cite{Abbott_2019}. The signals are generated using random right ascensions and
declinations uniform over the sky and the inclinations are drawn from the
cosine of a uniform distribution in the range [-1,1]. The peaks of the
waveforms are set to be within [0.4,0.6]s of the 1s time interval.

@@

% specifically define our architecture
%
We adopt the suggestions of \cite{Radford2015,DBLP:journals/corr/abs-1809-11096}, lengthening
one-dimensional convolution kernels on both the generator and discriminator.
The Generator model is fully convolutional, upsampled using strided transposed
convolutions with batch normalisation in the first layer and ReLU activations
throughout with the exception of Tanh for the output layer. Each transposed
convolutional layer uses a kernel size of $18\times 1$ and stride of 2. The
discriminator network mirrors that of the generator without batch
normalization, using LeakyReLU activations, SpatialDropout, and a 2-stride
convolution for downsampling. The discriminator has two output layers: the
first output is a single node activated by a Sigmoid that can be interpreted as
the realness of the the signal, the second output is 5 nodes activated using
the softmax function predicting the class of the input.  This model is trained
with binary cross entropy for the first output and sparse categorical
cross-entropy for the second output. The full architecture description can be seen in \cref{Tab:hyperparameters}.~\chris{OK, very technically tight but too
computer sciencey. Try to make it human readbale for a physicist and refer to
the table of parameters in the appendix.}

% describe the hyperparameter tunings
%
Neural networks and subsequently \acp{GAN} have multiple parameters a developer
can tune when designing the model and these are referred to as hyperparameters.
The final network design used in this work comes from the use of trial and
error and the initial designs influenced by the available literature. After
tuning the multiple hyperparamters (\cref{Tab:hyperparameters}), the GAN
was trained for 1000 epochs and takes $\Or$(2) days to train. ~\jordan{talk about making D and G 'symmetric' increasing batch size, number of filters powers of 2, spatial dropout,no early stopping.} ~\chris{trained on
what? Were there any major decisions made based on the trial and error process?
Anything that was learned from your mistakes that might be useful to others?}.

@@

% explain embedding
%
A learned embedding is an efficient way of representing categorical or class
based data. In contrast to traditional "one-hot encoding" where each class is
represented by a binary sparse vector, embeddings map each class to its own
distinct vector of a given size. The elements of these vectors are initialized
randomly and tuned during training just like weights~\chris{I thought that they
weren't trained. Are they definitely trained? Is this not just a single fully
connected layer that turns 5 numbers into 120? How is it any different to
that?}. ~\jordan{each class is one number that is turned into 120 numbers, embedding keeps track of the 5x120 numbers and as its training similar classes are grouped together in the class space. This explains it better than i can: https://machinelearningmastery.com/what-are-word-embeddings/ } This reduces computational cost as the sparse vectors containing mostly
zeros are not needed and allows related classes to cluster together. Once the
networks are trained the learned embedding vectors can be extracted from the
model and used as inputs to the generator. The benefit here is that the former
integer class labels are replaced by higher dimensional vector representations
allowing for finer experimenting in other applications. The class label is
interpreted as an additional channel early in the generator model. This is
achieved by projecting the class inputs as a learned embedding layer, into a
fully connected layer or "dense" layer which can then be concatenated
channel-wise to $\textbf{z}$.~\chris{very nice but still a bit too technical
without consideration for lowly astrophysicists. Can you soften it a bit to
make it more accessible. Explain more of the concepts when you introduce them.}

@@

\subsection{Waveform quality}
The generator network is a function G : $\mathbf{z},\mathbf{c},\mathbb{\textbf{s}}$ $\in$ $\mathbb{R}^{100}$ $\to$ $\mathbb{R}^{1024\times2}$, where $\mathbf{z},\mathbf{c},\mathbb{\textbf{s}}$ are the latent vector, class embedding vector and sky positions respectively. Given a latent vector randomly sampled from a normal distribution with zero mean and unit variance, a class label which is represented by a 120 dimensional vector for each class and antenna responses, the results from the generator can be seen in \cref{fig:gen_signals}. Depending on the orientation of the detector with respect to a hypothetical signal in the sky, the waveforms may appear inverted, shifted in time and their strain attenuated. Each plot shows the output of the generator after given randomised $\mathbf{z}, \mathbb{\textbf{s}}$ and one of the five class vectors $\mathbf{c}$.

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{figures/gens+taining.eps}
    \caption{Generated waveforms after training and conditioning on five classes. The generator will output two waveforms as seen by detectors in Hanford (red) and Livingston (blue). The generator is able to capture the characteristics of each waveform and structure the class space to give control over which waveform to generate. Each row shows a random assortment of one of the five classes the GAN is trained on. }
    \label{fig:gen_signals}
\end{figure}

@@

\subsection{Latent space interpolation}
In this section we explore the latent space formed by the generator by interpolation. We take two random points in the latent space and linearly interpolate between them (six times inclusive). These new latent space vectors can now be fed into the generator to make predictions on while keeping the class vectors constant. The third input, the response values are kept constant for each class. The full effect shown in \cref{fig:z_interp}. We can see that each plot shows plausible waveforms suggesting that the generator has constructed a smooth space unlike the discrete training case. Additionally as each class is given the same latent points to interpolate over, we can see that the waveforms cluster together with respect to their parameters. Visually, the sine-gaussian and ring-down waveforms share similar frequencies and the other signals show similar decays and starting epochs. The only exception is BBH waveforms, which is expected as they were trained with more variety of parameters and consistently have their peaks in the last quarter of the time series.

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{figures/z_interpolation.eps}
    \caption{Interpolations between two random latent points in z. Each row uniformly interpolates between two points in $\mathbf{z}$ keeping the class fixed. Only a single waveform from the generator is plotted and each signal is re-scaled to [-1,1] effectively removing the antenna responses for clarity.}
    \label{fig:z_interp}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/class_interpolation.eps}
    \caption{Class space interpolation with latent space held constant throughout. The plots show a zoomed in section of the 1s time interval. First row: Sine-Gaussian class to ringdown class, Second Row: Sine-Gaussian class to whitenoise burst class, Third row: Sine-Gaussian class to Gaussian class, Fourth row: Sine-Gaussian to BBH class.}
    \label{fig:c_interp}
\end{figure}
